---
title: "Analyst"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyst}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(opentripplanner)
```

The Analyst is a set of optional features in OTP 1.x used for transport analysis. These features were removed from OTP 2.0.  This Vignette will explain how to enable and use these features.

## Analyst Features

The Analys add the following feature to OTP:

1. **PointSets** - A more efficient way to specify a batch of fromPlace / toPlaces for routing.
1. **Surfaces** - A efficient way to get travel time from one to many places.
1. **Isochrones** - An alternative implementation of the Isochrone features.

## Making a PointSet and loading it into OTP

Any PointSets must be loaded when OTP starts, and so must be created before starting OTP. PointSets can be created from SF objects using the `otp_poinset()` function.

Using the same example data from the [Advanced Features Vignette](), the first steps are the same:

```{r eval=FALSE}
# Get OTP and the Demo Data
library(opentripplanner)
path_data <- file.path(tempdir(), "OTP")
dir.create(path_data)
path_otp <- otp_dl_jar(verion = "1.5.0") #Must use OTP 1.x
otp_dl_demo(path_data)
log1 <- otp_build_graph(otp = path_otp, dir = path_data)
```

Before starting OTP we need to create a pointset and put it in the correct location. We will use the demo LSOA points to create a PointSet. We will also add a column of data called jobs to analyse. For any numerical values in a PointSet OTP can provide the count and sum of that value based on travel criteria. In this case we migh want to know how many jobs can be accessed within a 30 minute drive of a given location.

```{r, eval=FALSE}
lsoa <- sf::st_read("https://github.com/ropensci/opentripplanner/releases/download/0.1/centroids.gpkg", stringsAsFactors = FALSE)
lsoa$jobs <- sample(100:500, nrow(lsoa))
otp_pointset(lsoa, "lsoa", path_data)
```

Here the `otp_pointset()` function takes the `lsoa` object and create a PointSet that can be accessed by OTP using the name "lsoa".

The next step is to start OTP, enable the analyst features, and load in the PointSets.

```{r eval=FALSE}
log2 <- otp_setup(otp = path_otp, dir = path_data, analyst = TRUE, pointsets = TRUE)
otpcon <- otp_connect(timezone = "Europe/London")
```

## Creating a Surface

Now we can create a surface and analyise out 


