---
title: "Dummies Guide"
author: "Malcolm Morgan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Introduction

This vignette is to provide guidance for people who are completely unfamiliar with Open Trip Planner and/or R.  It assumes no prior-knowledge.  If you are familiar with R and/or OTP then you may find the Getting Started vignette more useful.

## What is this package for, and what is a package anyway?

The Open Trip Planner for R package makes it easier for R and Open Trip Planner to communicate.

### What is Open Trip Planner?

OpenTripPlanner (OTP) is a free, open-source, and cross-platform multi-modal route planner written in JAVA.  It is like having your own private version of Google Maps.

You can find a lot more information about OTP at:

* http://www.opentripplanner.org/
* https://wiki.openstreetmap.org/wiki/OpenTripPlanner

### What is R?

R is a programming language and free software environment for statistical computing and graphics.  R has many capabilities for analysing data and writing software, but in this context, its ability to produce and analyse spatial data and maps is most relevant.

You can find a lot more information about R at:

* https://www.r-project.org/ 
* https://en.wikipedia.org/wiki/R_(programming_language) 

### What is an R package?

An R package is a small piece of software that extends the basic capabilities of R. It is a bit like how a new phone can do some things out of the box (make phone calls, send email) but you have to install apps to add extra abilities.

### Why would I want to use Open Trip Planner for R?

OTP can be used to run public route finding services such as https://ride.trimet.org but it can also be run on your own private computer or server. If you want, analyse a transport network OTP is a very useful and tool. However, OTP is, as the name suggests, a Trip Planner, not analytical software.  So, while OTP can find the fastest route from A to B or find all the places that are within a 20-minute walk of C, it cannot answer a question like “how many people live within 10-minutes of a park?”  This is where R can help.  R can process multiple spatial datasets such a population densities and park locations, but does not have a built-in journey planner. 

This package allows you to use the trip planning power of OTP with the analytical might of R.

## What do I need to get started?

You will need to install some software to use OTP and R.

### To use OTP

* Download free Java 8 from https://www.java.com/en/download/ and install Java. You will need the correct version for your operating system.
* Download the .shaded.jar OTP  file from https://repo1.maven.org/maven2/org/opentripplanner/otp/ at time of writing the latest version of OTP is version 1.3 so you would want to get https://repo1.maven.org/maven2/org/opentripplanner/otp/1.3.0/otp-1.3.0-shaded.jar 

For more information and advanced options, see http://docs.opentripplanner.org/en/latest/Getting-OTP/ 

For now just keep the .jar file somewhere where you can find it, we will use it later.

### To Use R

* Download free R software https://cran.r-project.org/mirrors.html selecting your country (or nearest available country)
* We strongly recommend you also download and install RStudio https://www.rstudio.com/products/rstudio/download/ that has a free version for personal use.

A video tutorial for installing R and RStudio 

* On Mac https://www.youtube.com/watch?v=cX532N_XLIs 
* On Windows https://www.youtube.com/watch?v=9-RrkJQQYqY 

### Learning R

If you have never used R before there a lots of free tutorials to get you started, such as:
* https://www.datacamp.com/courses/free-introduction-to-r
* http://www.r-tutor.com/r-introduction

### Getting the OTP for R package
Once you have R and RStudio installed open RStudio. There are two version of an R package the public release on CRAN and the development version on GitHub.

#### Getting the public CRAN version
**OTP for R is not yet available on CRAN, we are working towards a CRAN release soon.**

To get the latest release of OTP for R you will need to run the following lines of code.

```{r, eval = FALSE}
install.packages("opentripplanner")
```
When installing the OTP for R package, R will also install any other packages that are required. 

#### Getting the development version

The development version is a work in progress and may contain untested or incomplete features. **It is not recommended for beginners.** 
To get the latest development version you will need to run the following lines of code.

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("marcusyoung/opentripplanner")
```
This code will install a package called “devtools” and then use that package to install OTP for R from github. When installing the OTP for R package, R will also install any other packages that are required.

#### Starting the OTP for R Package

Once the package has been installed you will need to activate it by running this line of code.
```{r}
library(opentripplanner)
```

## Next Steps

Now that you have installed Open Trip Planner for R got to the Getting Started vignette, to find how to create a graph and use it to plan trips. You might also intrested in:

### Getting Data for OTP

To plan a trip OTP need data about the location, such as where the roads are, what is the public transport timetable etc.

#### Road Map Data

OTP uses road maps from the Open Street Map (OSM) https://www.openstreetmap.org in the .pbf format. OSM is a free map that anybody can edit. You can download all of part of the OSM from https://download.geofabrik.de/.

This package comes with some sample data for Greater Manchester https://download.geofabrik.de/europe/great-britain/england/greater-manchester.html

#### Public Transport Timetables

OTP can use public transport timetable data in the GFTS format. You can find GTFS data for many regions at https://transit.land/ 

#### Elevation Data

You can add terrain infromation to your routes, especially useful for walking and cycling using GeoTIFF images. You can find world wide elevation data at https://www2.jpl.nasa.gov/srtm/ .








