---
title: "OTP OpenTripPlanner Version 2.0"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OTPv2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `opentripplanner` package was originally developed to support OpenTripPlanner v1.3, and has been updated to support each subsequent release. OpenTripPlanner v2.0 is the next major release, and this vignette will document the changes within the package to support OpenTripPlanner v2.0. While most changes for v2.0 are improvements some, like the removal of the isochrone feature are not. So the package will continue to support both version 1.5 (the last v1.x) and subsequent v2.x releases. This vignette will be updated to help users select the best version for their needs.

**Major Changes in 2.0**

* Switch from Java 8 to Java 11
* Support for Netex transit data
* Support for SIRI transit data
* Switch from A* routing algorithm to Range Raptor when searching transit routes
* Removal of Isochrone, Geocode, and other analysis features

## Getting Java 11
At the time of writing Java 8 is still the default on may systems. To get Java 11 you must download the Java Development Kit https://www.oracle.com/java/technologies/javase-jdk11-downloads.html this requires a free account with the Oracle website.

On windows you will also need to change the PATH variable to point to the new version of Java

## Checking your Java version

You can check the version of Java accessible to R by using `otp_check_java` and specifying the version of OTP you want to use.

For example if you wanted to use OTP 2.0 but have Java 8 installed then you would see:

```{r, eval = FALSE}
otp_check_java(2)
[1] FALSE
Warning message:
In otp_check_java(2) : You have OTP 2.x but the version of Java for OTP 1.x

```

## Building a graph and starting up OTP 2

You can select OTP 2.0 using the `otp_dl_jar` function.


```{r setup, eval=FALSE}
library(opentripplanner)
# Path to a folder containing the OTP.jar file, change to where you saved the file.
path_data <- file.path(tempdir(), "OTP")
dir.create(path_data)
path_otp <- otp_dl_jar(version = "2.0.0")
otp_dl_demo(path_data)
# Build Graph and start OTP
log1 <- otp_build_graph(otp = path_otp, dir = path_data)
log2 <- otp_setup(otp = path_otp, dir = path_data)
otpcon <- otp_connect(timezone = "Europe/London")
```
We can download the demo data and build a graph in the usual way.



